---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# FLOC : Fast Limited-memory Optimal Change detector

<!-- badges: start -->
<!-- badges: end -->

The goal of FLOC is to sequentially detect a change in a piecewise linear 
model. 
The package also provides a method to tune the threshold of the algorithm to
bound the type I error, a function to estimate the false alarm probability and
a function to estimate the expected detection delay.

## Installation

You can install the development version of FLOC like so:

``` {r, eval = FALSE}
install.packages("devtools")
devtools::install_github("AnnikaHueselitz/FLOC")
```

## Example

This is a basic example which shows you how to solve a common problem:

For this example we want to let the algorithm run with bins that are 5
data points wide.

First we want to tune the type I error such that with 500 
historical observations available the probability to get a false alarm in the 
first 500 data points is approximately 0.5.

```{r thresholdtuning}
#load package
library(FLOC)

#set seed for reproducibility
set.seed(1)

#bin size
N <- 5

#amount of historical data
k <- 500

#the algorithm should have a false alarm in tau data points with a probability of eta
eta <- 0.5
tau <- 500 

#number of data sets to draw, the threshold tuning gets exacter when this is higher, but the computation is then slower
r <- 1000

#tuning of the threshold
rho <- t_tuning(N, tau, k, eta, r)
```

Next we want to simulate some data to test the algorithm on:

```{r detector}
#set seed for reproducibility
set.seed(2)

#true changepoint
tau_0 <- 200

#intercept at change point before change
alpha_minus <- 0

#slope before change
beta_minus <- 0

#intercept at change point after change
alpha_plus <- 1

#slope after change
beta_plus <- 0.1

#generating historical data
hist_data <- rnorm(k) + seg_lin_fun((1 - k):0, tau_0, alpha_minus, beta_minus, alpha_plus, beta_plus)

#initalize the detector
detector <- FLOC_init(hist_data, N, rho$both["jump"], rho$both["kink"])

#initialize time and vector to save data for plot
n <- 1
data <- c()

#generate new data as long as no detection was made and update the detector
while (!(detector$detect_jump||detector$detect_kink)) {
  
  #generate new data point
  new_data <- rnorm(1) + seg_lin_fun(n, tau_0, alpha_minus, beta_minus, alpha_plus, beta_plus)
  
  #save data to plot later
  data <- c(data, new_data)
  
  #update detector
  detector <- FLOC_iter(new_data, detector)
  
  n <- n+1
}

if (n <= tau_0) {
  print(paste("A detection was made at observation: ",n, "That is a false alarm"))
} else {
print(paste("A detection was made at observation: ",n, "That is a detection delay of: ", n - tau_0))
}

```
Plot the detection to visualize it:


```{r plot}

plot_detection(data, tau_0, alpha_minus, beta_minus, alpha_plus, beta_plus)

```


Next we can check if the false alarm probability that we wanted to achieve with the threshold tuning is reasonable, by estimating it:

```{r false alarm probability}
est_faprob(N, rho$both["jump"], rho$both["kink"], tau, k)
```

Last we can check estimate the expected detection delay for the above example:

```{r expected detection delay}
estimate_detdel(N, change_type, t, k, alpha_plus, beta_plus, alpha_minus, beta_minus)
```
